Usage metrics

### 2020-12-04 inst/analysis/usage_metrics.Rmd
Approximation of the number of runs per package

---
title: "Usage metrics"
author: "Pierre Donat-Bouillud"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR
  calls_packages_path: !r paste0(params$base_dir, "/run/package-evals-traced.5/summarized-packages.fst")
  corpus_path: !r paste0(params$base_dir, "/revalstudy/inst/data/corpus.fst")
---

@@

source("inc/paths.R")
source("inc/latextags.R")
source("inc/setup.R")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "enrichDataset.tex"), prefix="", default=TRUE)
```

@@

```{r read-data}
eval_calls <- read_fst(params$calls_packages_path) %>% as_tibble()
```

```{r read_corpus}
corpus <- read_fst(params$corpus_path) %>% as_tibble()
```

```{r eval-calls-overview}
own_eval_calls <- eval_calls %>% filter(eval_source == package) # package column is wrong currently
nb_eval_calls <- own_eval_calls %>% count(wt = nb_ev_calls) %>% select(n)

nb_eval_pres_calls <- own_eval_calls %>% filter(eval_function == "eval") %>% count(wt = nb_ev_calls) %>% select(n)
nb_evalq_calls <- own_eval_calls %>% filter(eval_function == "evalq") %>% count(wt = nb_ev_calls) %>% select(n)
nb_eval.parent_calls <- own_eval_calls %>% filter(eval_function == "eval.parent") %>% count(wt = nb_ev_calls) %>% select(n)
nb_local_calls <- own_eval_calls %>% filter(eval_function == "local") %>% count(wt = nb_ev_calls) %>% select(n)

@@

r("nb eval and variant calls", nb_eval_calls)
r("nb eval calls", nb_eval_pres_calls)
r("nb evalq calls", nb_evalq_calls)
r("nb eval parent calls", nb_eval.parent_calls)
r("nb local calls", nb_local_calls)
r("nb package eval at least once", eval_at_least_once)
```

@@

median_call_sites <- packages_at_least_one_call_site %>% summarise(med = median(nb_call_sites, na.rm = TRUE))
max_packages <- packages_at_least_one_call_site %>% filter(nb_call_site == max(nb_call_sites, na.rm = TRUE)) %>% select(eval_source, nb_call_site)

nb_no_calls_to_eval <- nrow(corpus) - nrow(packages_at_least_one_call_site)

@@

```{r call-sites-violin-plot}
call_sites_per_package %>%
 ggplot(aes(x = eval_source, y = nb_call_sites, fill = eval_function), na.rm = TRUE) +
  geom_violin(trim = FALSE, width=1.7)+
  geom_boxplot(data = slice_sample(nb_ev_call_sites, n = 100000), width=.03) + # too slow otherwise
  #scale_y_log10(labels = scales::comma) +
  labs(y="Call sites per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background = element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

ggsave(path(PLOT_DIR, "call_sites_per_packages.pdf"))
```

@@

This is the number of eval calls which have a call site in each package.

```{r nb-eval-calls}
# run = package?
eval_calls_per_package <- own_eval_calls %>% count(eval_source, wt=nb_ev_calls)

average_eval_calls <- eval_calls_per_package %>% filter(n > 0) %>% summarise(, average = mean(n)) %>% .["average"]
max_eval_calls <- eval_calls_per_package %>% filter(n = max(n))

r("average eval calls per package", average_eval_calls) # average on packages with at least one eval call
r("max eval calls", max_eval_calls$n)
r("max eval call package", max_eval_calls$eval_source)
```

```{r eval-calls-violin-plot}
call_sites_per_package %>%
 ggplot(aes(x = eval_source, y = n), na.rm = TRUE) +
  geom_violin(trim = FALSE, width=1.7)+
  geom_boxplot(data = slice_sample(nb_ev_call_sites, n = 100000), width=.03) + # too slow otherwise
  #scale_y_log10(labels = scales::comma) +
  labs(y="Eval calls per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background = element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

ggsave(path(PLOT_DIR, "call_sites_per_packages.pdf"))
```

```{r}
nb_eval_calls_tot <- eval_calls %>% count(wt = nb_ev_calls) %>% select(n)
```

If we don't exclude dependent calls, there are `r nb_eval_calls_tot` eval calls, so the proper eval calls are `r percent(nb_eval_calls, nb_eval_calls_tot)` of the total.

@@

med_size <- own_eval_calls %>% summarise(med = weighted.median(expr_expression_length, na.rm = TRUE, w = nb_ev_calls))
max_size <- own_eval_calls %>% summarize(m = max(expr_expression_length, na.rm = TRUE))
average_size <- own_eval_calls %>% summarize(average = weithed.mean(expr_expression_length, na.rm = TRUE, w = nb_eval_calls))

sizes_prop <- own_eval_calls %>% count(expr_expression_size, wt = nb_ev_calls) %>%
                      mutate(proportion = n / sum(n)) %>% 
                      arrange(expr_expression_size) %>% 
                      mutate(cumulative = cumsum(proportion))

@@

r("size nine five", nb_nodes_many$expr_expression_size)
r("size nine five exact percent", percent(nb_nodes_many$cumulative))

@@

r("nb events eval", nb_events_in_eval$nb_events)
r("maxi events eval", max_events_in_eval$maxi)
r("small events eval", percent(count(nb_less_100_events), nb_eval_calls))
```

```{r nb-events-violin-plot}
own_eval_calls %>%
 ggplot(aes(x = eval_source, y = direct_interpreter_eval), na.rm = TRUE) +
  geom_violin(trim = FALSE, width=1.7)+
  geom_boxplot(data = slice_sample(direct_interpreter_eval, n = 100000), width=.03) + # too slow otherwise
  #scale_y_log10(labels = scales::comma) +
  labs(y="Number of events in eval per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background = element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

>>>>

---
title: "Usage metrics"
author: "Pierre Donat-Bouillud"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR
  calls_packages_path: /run/package-evals-traced.7/summarized-externals.fst
  corpus_path: /revalstudy/inst/data/corpus.fst
---

@@

source("inc/paths.R")
source("inc/latextags.R")
source("inc/setup.R")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "usage_metrics.tex"), prefix="", default=TRUE)
```

@@

```{r read-data}
eval_calls <- read_fst(paste0(params$base_dir, params$calls_packages_path)) %>% as_tibble()
```

```{r read_corpus}
corpus <- read_fst(paste0(params$base_dir, params$corpus_path)) %>% as_tibble()
```

```{r eval-calls-overview}
#own_eval_calls <- eval_calls %>% filter(eval_source == package) # probably to remove
own_eval_calls <- eval_calls
nb_eval_calls <- (own_eval_calls %>% count(wt = nb_ev_calls))$n

nb_eval_pres_calls <- own_eval_calls %>% filter(eval_function == "eval") %>% count(wt = nb_ev_calls)
nb_evalq_calls <- own_eval_calls %>% filter(eval_function == "evalq") %>% count(wt = nb_ev_calls)
nb_eval.parent_calls <- own_eval_calls %>% filter(eval_function == "eval.parent") %>% count(wt = nb_ev_calls)
nb_local_calls <- own_eval_calls %>% filter(eval_function == "local") %>% count(wt = nb_ev_calls)

@@

r("nb eval and variant calls", nb_eval_calls)
r("nb eval calls", nb_eval_pres_calls$n)
r("nb evalq calls", nb_evalq_calls$n)
r("nb eval parent calls", nb_eval.parent_calls$n)
r("nb local calls", nb_local_calls$n)
r("nb package eval at least once", eval_at_least_once)
```

@@

median_call_sites <- packages_at_least_one_call_site %>% summarise(med = median(nb_call_sites, na.rm = TRUE))
max_packages <- packages_at_least_one_call_site %>% filter(nb_call_sites == max(nb_call_sites, na.rm = TRUE)) %>% select(eval_source, nb_call_sites)

nb_no_calls_to_eval <- nrow(corpus) - nrow(packages_at_least_one_call_site) # corpus seems to be wrong

@@

```{r call-sites-violin-plot}
call_sites_per_package %>%
 ggplot(aes(x = eval_source, y = nb_call_sites, fill = eval_function), na.rm = TRUE) +
  geom_violin(trim = FALSE, width=1.7)+
  geom_boxplot(data = slice_sample(call_sites_per_package, n = 100000), width=.03) + # too slow otherwise
  #scale_y_log10(labels = scales::comma) +
  labs(y="Call sites per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background = element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

#ggsave(path(PLOT_DIR, "call_sites_per_packages.pdf"))
```

@@

This is the number of eval calls which have a call site in each package, divided by the number of runs for the package.

We first compute an approximation of the number of runs per package.
```{r runs-per-package}
runs_package <- own_eval_calls %>% group_by(eval_source) %>% summarise(nb_runs = n_distinct(file)) %>% arrange(desc(nb_runs))
```

What are the packages with the most runs?

```{r show_run-per-package}
runs_package %>% datatable()
```

```{r nb-eval-calls}
# run = package?
eval_calls_per_package <- own_eval_calls %>% count(eval_source, wt=nb_ev_calls) %>% left_join(runs_package, by = "eval_source") %>% mutate(n = n / nb_runs)

average_eval_calls <- eval_calls_per_package %>% filter(n > 0) %>% summarise(average = mean(n))
max_eval_calls <- eval_calls_per_package %>% filter(n == max(n))

r("average eval calls per package", average_eval_calls$average) # average on packages with at least one eval call
r("max eval calls", max_eval_calls$n)
r("max eval call package", max_eval_calls$eval_source)
```

```{r eval-calls-violin-plot}
eval_calls_per_package %>%
 ggplot(aes(x = eval_source, y = n), na.rm = TRUE) +
  geom_violin(trim = FALSE, width=1.7)+
  geom_boxplot(data = slice_sample(eval_calls_per_package, n = 100000), width=.03) + # too slow otherwise
  #scale_y_log10(labels = scales::comma) +
  labs(y="Eval calls per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background = element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

ggsave(path(PLOT_DIR, "eval_calls_per_packages.pdf"))
```

```{r}
nb_eval_calls_tot <- eval_calls %>% count(wt = nb_ev_calls)
```

If we don't exclude dependent calls, there are `r nb_eval_calls_tot` eval calls, so the proper eval calls are `r ratio(nb_eval_calls, nb_eval_calls_tot$n)` of the total.

@@

weighted.median <- function(x, w) {
  w <- w[order(x)]
  x <- x[order(x)]

  prob <- cumsum(w)/sum(w)
  ps <- which(abs(prob - .5) == min(abs(prob - .5)))
  return(x[ps])
}

med_size <- own_eval_calls %>% summarise(med = median(expr_expression_length, na.rm = TRUE)) # should be weighted by nb_ev_calls
max_size <- own_eval_calls %>% summarize(m = max(expr_expression_length, na.rm = TRUE))
average_size <- weighted.mean(own_eval_calls$expr_expression_length, na.rm = TRUE, w = own_eval_calls$nb_ev_calls)

sizes_prop <- own_eval_calls %>% count(expr_expression_length, wt = nb_ev_calls) %>%
                      mutate(proportion = n / sum(n)) %>% 
                      arrange(expr_expression_length) %>% 
                      mutate(cumulative = cumsum(proportion))

@@

r("median size eval", med_size)
r("max size eval", max_size)
r("average size eval", average_size)
r("size nine five", nb_nodes_many$expr_expression_length)
r("size nine five exact percent", percent(nb_nodes_many$cumulative))

@@

r("nb events eval", nb_events_in_eval$nb_events)
r("maxi events eval", max_events_in_eval$maxi)
r("small events eval", ratio(nb_less_100_events, nb_eval_calls))
```

```{r nb-events-violin-plot}
own_eval_calls %>%
 ggplot(aes(x = eval_source, y = direct_interpreter_eval), na.rm = TRUE) +
  geom_violin(trim = FALSE, width=1.7)+
  geom_boxplot(data = slice_sample(own_eval_calls, n = 100000), width=.03) + # too slow otherwise
  #scale_y_log10(labels = scales::comma) +
  labs(y="Number of events in eval per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background = element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )
